### 2019-05-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Chapter 6: talk about [Luo et al, COLT, 2018]

@@

Three recent works proposed algorithms that achieve a small regret under the two models.
For simple bandits, \cite{WeiSrivastava18Abruptly} proposed the SW-UCB\# and LM-DSEE algorithms FIXME.
For contextual bandits, \cite{Luo18} and \cite{ChenLeeLuoWei2019} FIXME.
\TODOL{Parler de ces travaux là, qui sont très forts mais pas non plus encore parfaits.}

@@

\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
%
In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
we assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
%
This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.

% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}

@@

Then, we compare our algorithms to several competitor designed for a piece-wise stationary model. For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}. Moreover, all the algorithms are tuned as described in the corresponding paper, using in particular the knowledge of the number of breakpoints $\Upsilon_T$ and the horizon $T$. We first include three \emph{passively adaptive algorithms}:
Discounted \klUCB{} (D-\klUCB, \cite{Kocsis06}), with discount factor $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$; Sliding-Window \klUCB{} (SW-\klUCB,  \cite{Garivier11UCBDiscount}) using window-size $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$ and Discounted Thompson sampling (DTS, \cite{RajKalyani17}) with discount factor $\gamma = 0.95$. For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on the problem instances we tried.
More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).

% \TODOE{Tu confirmes bien que $\gamma=0.75$ ne marchait sur AUCUN problème ? Car ta phrase actuelle n'exclut pas que ce choix marche pour les probèmes courts}

@@

The two Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2} show the final regret $R_T$ obtained for each algorithm.
Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiments,
with our proposal being the best non-oracle strategy for problems $1$ and $2$.
Thompson sampling and \klUCB{} are efficient,
and better than Discounted-\klUCB{} which is very inefficient.
DTS and SW-\klUCB{} can sometimes be more efficient than their stationary counterparts, but perform worse than the Oracle and most actively adaptive algorithms.
M-\klUCB{} and \CUSUMklUCB{} outperform the previous algorithms, but \GLRklUCB{} is often better.
%
On these problems, \GLRklUCB{} with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts. Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.

\TODOL{I could add the following algorithms, just to illustrate more:

% - OracleRestart with UCB
% - M- and CUSUM- with UCB (as defined in the papers)
- Exp3.S
}

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|cccccc}
    \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $1$ & Pb $2$ & Pb $3$ \\
        \hline
        Oracle-Restart \klUCB{} & $\mathbf{37 \pm 37}$ & $\mathbf{45 \pm 34}$ & $\mathbf{257 \pm 86}$ \\
        \hline
        \klUCB{} & $270 \pm 76$ & $162 \pm 59$ & $529 \pm 148$ \\
        Discounted-\klUCB{} & $1456 \pm 214$ & $1442 \pm 440$ & $1376 \pm 37$ \\
        SW-\klUCB{} & $177 \pm 34$ & $182 \pm 34$ & $1794 \pm 71$ \\
        \hline
        Thompson sampling & $493 \pm 175$ & $388 \pm 147$ & $1019 \pm 245$ \\
        DTS & $209 \pm 38$ & $249 \pm 39$ & $2492 \pm 52$ \\
        \hline
        \MklUCB{} & $290 \pm 29$ & $534 \pm 93$ & $645 \pm 141$ \\
        \CUSUMklUCB{} & $148 \pm 32$ & $152 \pm 42$ & $\mathbf{490 \pm 133}$ \\
        \hline
        \GLRklUCB{}(Local) & $\mathbf{74 \pm 31}$ & $\mathbf{113 \pm 34}$ & $513 \pm 97$ \\
        \GLRklUCB{}(Global) & $97 \pm 32$ & $134 \pm 33$ & $621 \pm 103$
    \end{tabular}
    \caption{Mean regret $\pm$ $1$ std-dev, for different algorithms on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).
    }
    \label{table:6:totalResults1}
% \end{footnotesize}
\end{table}

\begin{table}[ht]
    % \begin{small}
    \centering
    \begin{tabular}{c|cccccc}
        \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $4$ ($T=5000$) & Pb $4$ ($T=10000$) & Pb $5$ \\
        \hline
        Oracle-Restart \klUCB{} & $\mathbf{68 \pm 40}$ & $\mathbf{86 \pm 50}$ & $126 \pm 54$ \\
        \hline
        \klUCB{} & $615 \pm 74$ & $1218 \pm 123$ & $106 \pm 36$ \\
        SW-\klUCB{} & $202 \pm 33$ & $322 \pm 47$ & $228 \pm 27$ \\
        Discounted-\klUCB{} & $911 \pm 210$ & $1741 \pm 200$ & $2085 \pm 910$ \\
        \hline
        Thompson sampling & $756 \pm 65$ & $1476 \pm 137$ & $\mathbf{88 \pm 39}$ \\
        DTS & $250 \pm 39$ & $481 \pm 58$ & $238 \pm 24$ \\
        \hline
        \MklUCB{} & $337 \pm 46$ & $544 \pm 47$ & $116 \pm 36$ \\
        \CUSUMklUCB{} & $267 \pm 69$ & $343 \pm 94$ & $117 \pm 34$ \\
        \hline
        \GLRklUCB{}(Local) & $\mathbf{99 \pm 32}$ & $\mathbf{128 \pm 42}$ & $149 \pm 34$ \\
        \GLRklUCB{}(Global) & $128 \pm 32$ & $185 \pm 47$ & $152 \pm 32$
    \end{tabular}
    \caption{Mean regret $\pm$ $1$ std-dev. Problem $4$ use $K=3$ arms, and a first long stationary sequence. Problem $5$ use $K=5$, $T=20000$ and is much harder with $\Upsilon=82$ breakpoints and $C=179$ changes.}
    \label{table:6:totalResults2}
    % \end{small}
\end{table}

@@

%----------------------------------------------------------------------------
\section{Conclusion}
\label{sub:6:conclusion}

>>>>

Another interesting recent work is \cite{AuerGajaneOrtner18}, where the authors propose an efficient algorithm, AdSwitch, for the two-armed case ($K=2$).
It proceeds in episodes, starting a new episode when the algorithm detects a change in one of the arms. Each episode is an estimation phase, playing both arms alternatingly until their means can be distinguished, then an exploitation phase. With some low probability, an exploration phase is started that checks whether a change has occurred.
Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that algorithm does not need any prior knowledge of the minimum magnitude.
Instead, it uses geometrically decreasing magnitudes $d_i = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only consider a few different values of $d_i$.
% Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.  Each episode startswith an estimation phase (Step 2), in which both arms are chosen alternatingly until theirmeans can be distinguished (Step 3).  After the estimation phase,  the algorithm exploitsthe empirical best arm (Step 5).  However,  in order to detect changes,  at each step withsome probability (Step 4) an exploration phase is started that checks whether a change hasoccurred (comparing the resulting empirical means to that obtained from the estimationphase).  During such exploration phases the algorithm checks for changes of certain magni-tudes (denoted bydi).  Note that our algorithm does not require knowledge of the numberof changesL.  The following theorem provides an upper bound on the expected regret ofour algorithm
They prove it achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
However, this work as two drawbacks:
first, the algorithm currently only applies to two arms, and the generalization does not seem straightforward.
Then, the theoretical result is valid for ``sufficiently large constants $C_1$ and $C_2$'', with a large constant $C$ hiding in the $\cO(\bullet)$ notation.
We did some experiments with AdSwitch, even if it is not included in the benchmark presented below in Section~\ref{sec:6:NumericalExperiments}. It uses two parameters $C_1,C_2$ that are set at large values for the theoretical analysis,
and while we explored numerically different choices, it seemed that setting $C_1=C_2=1$ gave the best performance (which is outside of the comfort zone of the theoretical analysis). Even for this tuning, and for problems with just $K=2$ arms, empirically   this policy is performing poorly in comparison to other policies based on active CD detection and \klUCB{} indexes.
%  We did not find a practical tuning outperforming GLR-klUCB yet.
% It would be interesting to compare

@@

@@

\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
%
In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
we assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
%
This interpretation is interesting too,
% but due to time and space constraints, we preferred to not explore it further.
but empirically we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
%
% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}

\paragraph{Adversarial or non-stationary contextual bandits.}
%
For contextual bandits, two very recent works are \cite{Luo18} and \cite{ChenLeeLuoWei2019}.
While both works target a more general setting, their algorithms are also applicable to non-contextual bandits, \ie, classical bandits like the model in this chapter.
From Table~1 in \cite{Luo18}, one can see that their two algorithms Exp4.S and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
that is, a $\cO(\Upsilon_T \sqrt{T}$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
We discuss both these algorithms:
%
first, Exp4.S actually reduces to Exp3.S in the non-contextual case, and Exp3.S is discussed above.
% As already acknowledged in the answer to  Reviewer 1,  we were indeed unaware of the very good theoretical properties of this algorithm in the piece-wise non-stationary setting. We will revise our state-of-the-art section and include the Exp3.S algorithm as a baseline in our experiments. However, preliminary experiments reveal that this algorithm is much less efficient that GLR-klUCB.
%
Then, the Ada-ILTCB algorithm is however a less appealing candidate for our setting, for the following reason.
This algorithm requires as input a parameter $L$ which needs to be an upper bound on the largest stationary sequence in the problem, in order for its regret to scale as $\Upsilon_T \sqrt{L}$ (neglecting $K$ and $\log$ factors).
It is claimed that the tuning $L=T/\Upsilon_T$ yields the optimal bound $\sqrt{\Upsilon_T T}$.
However, this tuning is only possible in a ``balanced'' instance without stationary sequence longer than $T/\Upsilon_T$.
In a piece-wise stationary instance with a stationary sequence of length $T/2$, one cannot get better than $\cO(\Upsilon_T \sqrt{T})$ regret, by setting $L=T/2$ (see for instance the problem 4 in our benchmark below in Section~\ref{sec:6:NumericalExperiments}), and as such the regret bound of Ada-ILTCB appear much worse than the announced $\cO(\sqrt{\Upsilon_T T})$ bound.
%
Besides, no experiments are reported by \cite{Luo18}, and the Ada-ILTCB algorithm seems much more complicated to implement than other alternatives,
so we prefer to not include it in the experiments presented later.
% with a valid tuning of L in each case.

\TODOL{Okay ça suffit pour ces travaux ?}

@@

Then, we compare our algorithms to several competitor designed for a piece-wise stationary model. For a fair comparison, all algorithms that use  \UCB{} as a sub-routine were adapted to use \klUCB{} instead, which yields better performance\footnote{\cite{LiuLeeShroff17,CaoZhenKvetonXie18} both mention that extending their analysis to the use of \klUCB{} should not be too difficult.}. Moreover, all the algorithms are tuned as described in the corresponding paper, using in particular the knowledge of the number of breakpoints $\Upsilon_T$ and the horizon $T$. We first include three \emph{passively adaptive algorithms}:
Discounted \klUCB{} (D-\klUCB, \cite{Kocsis06}), with discount factor $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$; Sliding-Window \klUCB{} (SW-\klUCB,  \cite{Garivier11UCBDiscount}) using window-size $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$ and Discounted Thompson sampling (DTS, \cite{RajKalyani17}) with discount factor $\gamma = 0.95$. For this last algorithm, the discount factor $\gamma=0.75$ suggested by the authors was performing significantly worse on the problem instances we tried.
More precisely, we found that $\gamma\leq0.95$ gives better performances for short-term problems (pbs $1,2,4$) and $\gamma\geq0.95$ is better suited for long experiments (pbs $3,5$).
%
% \TODOE{Tu confirmes bien que $\gamma=0.75$ ne marchait sur AUCUN problème ? Car ta phrase actuelle n'exclut pas que ce choix marche pour les probèmes courts}
%
Additionally, we include the Exp3.S algorithm from \cite{Auer02NonStochastic}, setting its parameters as in Corollary 8.3 of the paper, based on a prior knowledge of $T$ and $\Upsilon_T$, using $\alpha=1/T$ and $\gamma = \min(1, \sqrt{K (\Upsilon_T \log(KT) + \mathrm{e}) / ((\mathrm{e}-1)T)})$.

@@

The two Tables~\ref{table:6:totalResults1} and \ref{table:6:totalResults2} show the final regret $R_T$ obtained for each algorithm.
Results highlighted in \textbf{bold} show the best non-oracle algorithm for each experiments,
with our proposal being the best non-oracle strategy for problems $1$ and $2$.
Thompson sampling and \klUCB{} are efficient,
and better than Discounted-\klUCB{} which is very inefficient.
DTS and SW-\klUCB{} can sometimes be more efficient than their stationary counterparts, but perform worse than the Oracle and most actively adaptive algorithms.
M-\klUCB{} and \CUSUMklUCB{} outperform the previous algorithms, but \GLRklUCB{} is often better.
%
On these problems, \GLRklUCB{} with \textbf{Local} restarts is always more efficient than with \textbf{Global} restarts. Note that on problem 2, all means change at every breakpoint, hence one could expect global restart to be more efficient, yet the experiments show the superiority of local restarts on every instances.
The Exp3.S algorithm was found to outperform other algorithms based on Exp3, including Exp3, and recent variants like Exp3$++$ \cite{Seldin17} or Exp3.R from \cite{Allesiardo17},
and it usually performs similarly to \MklUCB, however it is greatly outperformed by the oracle algorithm, by \GLRklUCB{} and by \CUSUMklUCB.

\TODOL{I HAVE TO add the following algorithms, just to illustrate more Exp3.S}

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|cccccc}
    \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $1$ & Pb $2$ & Pb $3$ \\
        \hline
        Oracle-Restart \klUCB{} & $\mathbf{37 \pm 37}$ & $\mathbf{45 \pm 34}$ & $\mathbf{257 \pm 86}$ \\
        \hline
        Exp3.S & $XXX$ & $XXX$ & $XXX$ \\
        \hline
        \klUCB{} & $270 \pm 76$ & $162 \pm 59$ & $529 \pm 148$ \\
        Discounted-\klUCB{} & $1456 \pm 214$ & $1442 \pm 440$ & $1376 \pm 37$ \\
        SW-\klUCB{} & $177 \pm 34$ & $182 \pm 34$ & $1794 \pm 71$ \\
        \hline
        Thompson sampling & $493 \pm 175$ & $388 \pm 147$ & $1019 \pm 245$ \\
        DTS & $209 \pm 38$ & $249 \pm 39$ & $2492 \pm 52$ \\
        \hline
        \MklUCB{} & $290 \pm 29$ & $534 \pm 93$ & $645 \pm 141$ \\
        \CUSUMklUCB{} & $148 \pm 32$ & $152 \pm 42$ & $\mathbf{490 \pm 133}$ \\
        \hline
        \GLRklUCB{}(Local) & $\mathbf{74 \pm 31}$ & $\mathbf{113 \pm 34}$ & $513 \pm 97$ \\
        \GLRklUCB{}(Global) & $97 \pm 32$ & $134 \pm 33$ & $621 \pm 103$
    \end{tabular}
    \caption{Mean regret $\pm$ $1$ std-dev, on problems $1$, $2$ (with $T=5000$) and $3$ ($T=20000$).
    }
    \label{table:6:totalResults1}
% \end{footnotesize}
\end{table}

\begin{table}[ht]
    % \begin{small}
    \centering
    \begin{tabular}{c|cccccc}
        \textbf{Algorithms} $\;$ \textbackslash $\;$ \textbf{Problems} & Pb $4$ ($T=5000$) & Pb $4$ ($T=10000$) & Pb $5$ \\
        \hline
        Oracle-Restart \klUCB{} & $\mathbf{68 \pm 40}$ & $\mathbf{86 \pm 50}$ & $126 \pm 54$ \\
        \hline
        Exp3.S & $XXX$ & $XXX$ & $XXX$ \\
        \hline
        \klUCB{} & $615 \pm 74$ & $1218 \pm 123$ & $106 \pm 36$ \\
        SW-\klUCB{} & $202 \pm 33$ & $322 \pm 47$ & $228 \pm 27$ \\
        Discounted-\klUCB{} & $911 \pm 210$ & $1741 \pm 200$ & $2085 \pm 910$ \\
        \hline
        Thompson sampling & $756 \pm 65$ & $1476 \pm 137$ & $\mathbf{88 \pm 39}$ \\
        DTS & $250 \pm 39$ & $481 \pm 58$ & $238 \pm 24$ \\
        \hline
        \MklUCB{} & $337 \pm 46$ & $544 \pm 47$ & $116 \pm 36$ \\
        \CUSUMklUCB{} & $267 \pm 69$ & $343 \pm 94$ & $117 \pm 34$ \\
        \hline
        \GLRklUCB{}(Local) & $\mathbf{99 \pm 32}$ & $\mathbf{128 \pm 42}$ & $149 \pm 34$ \\
        \GLRklUCB{}(Global) & $128 \pm 32$ & $185 \pm 47$ & $152 \pm 32$
    \end{tabular}
    \caption{Mean regret $\pm$ $1$ std-dev. Problem $4$ use $K=3$ arms, and a first long stationary sequence. Problem $5$ use $K=5$, $T=20000$ and is much harder with $\Upsilon=82$ breakpoints and $C=179$ changes.}
    \label{table:6:totalResults2}
    % \end{small}
\end{table}

@@

\newpage
%----------------------------------------------------------------------------
\section{Conclusion}
\label{sub:6:conclusion}
